\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-05-22 manuscript.tex
comments, minor edits

\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; QTL \lev{??}

@@

We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and phased data from \citet{Pyhajarvi2013}, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}. SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}. We created haplotype blocks using a custom Perl script \lev{is this where I should cite Matt's paper, or his code? Matt is was your 'chunkfy.pl script} that grouped SNPs separated by less than 5kb into haplotypes. We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.

\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1). We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set. In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion. We chose between 10-13 seeds from each of 23 sampling sites. We treated seeds with Captan fungicide (Southern Agricultural Insecticides, Inc.) \lev{do we need to say which?} and germinated them in petri dishes with filter paper. Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches. Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer.

@@

Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; \lev{alpha is a parameter used in Box-Cox transformations} Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.  We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA). Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.

@@

\textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals. Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2). Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig1Map}). 
%We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population. Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p$<$0.05). \mbh{Jeff mentioned this needs to be fully reported in a supplementary table; alternatively we can just delete the result since we do not interpret or use it later in the manuscript.}
\lev{so are you wanting to remove the above blip about hew stuff?}

>>>>

\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; QTL \lev{??}  \mbh{not sure about QTL...might make people think we've done some QTL mapping which we haven't}

@@

We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and phased data from \citet{Pyhajarvi2013}, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}. SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}. We created haplotype blocks using a custom Perl script \lev{is this where I should cite Matt's paper, or his code? Matt is was your 'chunkfy.pl script} \mbh{we should make this script available; my vote would be to get this submitted and during review get all scripts, etc... in GitHub or some other public repository} that grouped SNPs separated by less than 5kb into haplotypes. We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.

\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1). We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set. In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion. We chose between 10-13 seeds from each of 23 sampling sites. We treated seeds with Captan fungicide (Southern Agricultural Insecticides Inc., Palmetto, Florida, USA) and germinated them in petri dishes with filter paper. Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches. Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer.

@@

Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; \lev{alpha is a parameter used in Box-Cox transformations} \mbh{the box-cox transformation function in the car package uses a lambda vector, but I do not see mention of alpha} Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.  We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA). Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.

@@

\textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals. Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2). Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig1Map}). 
%We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population. Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p$<$0.05). \mbh{Jeff mentioned this needs to be fully reported in a supplementary table; alternatively we can just delete the result since we do not interpret or use it later in the manuscript.}
\lev{so are you wanting to remove the above blip about hew stuff?}
\mbh{I say ditch it, we never come back to it}
